---
title: "Assignment/HW5"
author: "Group 2"
date: "April 19, 2018"
output: html_document
---
Overview
In this homework assignment we explore, analyze, and model a data set containing information on approximately 12,000 commercially available wines. The variables are mostly related to the chemical properties of the wine being sold. The response variable is the number of sample cases of wine that were purchased by wine distribution companies after sampling a wine. These cases would be used to provide tasting samples to restaurants and wine stores around the United States. The more sample cases purchased, the more likely is a wine to be sold at a high end restaurant. 

A large wine manufacturer is studying the data in order to predict the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then it will be able to adjust their wine offering to maximize sales.

Our objective is to build a count regression model to predict the number of cases of wine that will be sold
given certain properties of the wine. 
```{r message=FALSE, warning=FALSE}
library(psych)
library(GGally)
#library(ggplot2)
library(reshape)
library(VIM)
library(mice)
#library(stringr)
#library(dplyr)
library(car)
library(usdm)
library(tidyverse)
#library(stringr)
library(DataExplorer)
library(knitr)
library(corrplot)
library(MASS)
#library(tinytex)
library(ggfortify)
library(caret)
library(pROC)
library(robust)
library(robustbase)
library(pscl)
library(MKmisc)
library(Metrics)  # for lm metric calculation
library(gvlma)  ## only used for confirming model assumptions

options(scipen=999)
```
#1. DATA EXPLORATION

The data were loaded into a Github repository for better reproducability. After taking a quick look at the head of the data, there are a number of `NA`'s we will have to deal with. We will also remove the `INDEX` column as it will not be useful in our analysis. 

```{r}
wine_data <- read.csv("https://raw.githubusercontent.com/621-Group2/HW5/master/wine-training-data.csv")
head(wine_data)
wine_data$INDEX <- NULL
```

```{r}
# get_outliers function
get_outliers <-  function(x, n = 10) {
  
  bp <- boxplot.stats(x)
  
  obs_hi <- unique(x[which(x > bp$stats[5])])

  if (length(obs_hi) < n) 
    
    { n <- length(obs_hi) }

    hi <- sort(obs_hi, decreasing = T)[1:n]
  
    obs_low <- unique(x[which(x < bp$stats[1])])

  if (length(obs_low) < n) 
    
    { n <- length(obs_low) }

    low <- sort(obs_low, decreasing = T)[1:n]

  return (list(Hi=hi, Low=low))
  
}  
```


##Data Exploration and Statistic Measures

The purpose of the data exploration and statistic measures phase is to understand the data to determine how to process the dataset for modelling.
```{r}
variable_names <- c("Target", "AcidIndex", "Alcohol", "Chlorides", 
                    "CitricAcid", "Density", "FixedAcidity", "FreeSulfurDioxide", "LabelAppeal", 
                    "ResidualSugar", "Stars", "Sulphates", "TotalSulfurDioxide", "VolatileAcidity", "pH")

description <- c(  "Number of cases purchased", 
                   "Proprietary method of testing total acidity of wine by using a weighted average", 
                   "Alcohol content", 
                   "Chloride content of wine",
                   "Citric acid content",
                   "Density of wine",
                   "Fixed acidity of wine",
                   "Sulfur dioxide content of wine",
                   "Marketing score indidcating appeal of label design for consumers. Positive is like, negative is dislike",
                   "Resdiual sugar of wine",
                   "Wine rating by a team of experts. 4 stars = Excellent, 1 star = Poor",
                   "Sulfate content of wine",
                   "Total sulfur dioxide of wine",
                   "Volatile acid content of wine",
                   "pH of wine")

variable_type <-c("Response", "Predictor", "Predictor", "Predictor", "Predictor", 
                  "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", 
                  "Predictor", "Predictor", "Predictor",  "Predictor",  "Predictor")

#data_type <- c("quantitative", "quantitative", "quantitative", "quantitative", "quantitative", 
#               "quantitative", "quantitative", "quantitative", "categorical", "quantitative", 
#               "categorical", "quantitative", "quantitative", "quantitative", "quantitative")

wine_descriptions <- cbind.data.frame(variable_names, description, variable_type)

colnames(wine_descriptions) <- c("Variable Name", "Description", "Variable Type") 

kable(wine_descriptions)
```


The training set has 12,795 observations on 1 response and 14 predictor variables, representing certain charachteristics about a bottle of wine.
```{r}
glimpse(wine_data)
```

##Descriptive Statistics

Descriptive statistics were performed for all predictor and response variables to explore the data. Some of our predictors are numerical continuous (such as `VolatileAcidity` or `Density`), while others are numerical discrete (such as `LabelAppeal` or `TARGET`). There are no categorical variables in this data set.

We will now look at the summary of descriptive statistics:
```{r}
#Calculate mean missing values per variable
missing_values <- wine_data %>% 
  summarize_all(funs(sum(is.na(.))))

missing_values_ratio <- wine_data %>% 
  summarize_all(funs(sum(is.na(.)) / length(.)*100))

#Use Describe Package to calculate Descriptive Statistic
(wine_data_des <- describe(wine_data, na.rm=TRUE, interp=FALSE, skew=TRUE, ranges=TRUE, trim=.1, type=3, check=TRUE, fast=FALSE, quant=c(.25,.75), IQR=TRUE))

wine_data_des$missing <- t(missing_values)
wine_data_des$miss_ratio <- t(round(missing_values_ratio,4))

wine_data_des_display <- wine_data_des %>% 
  dplyr::select(n, missing, miss_ratio, mean, sd, min, max, skew, kurtosis, median, IQR, Q0.25, Q0.75)

knitr::kable(wine_data_des_display)
```


###Analysis of predictors

We will now examined each predictor to understand their distribution and determine whether any transformation is required.
Looking at the skewness coefficient and kurtosis, `AcidIndex` appears to be most skewed.
```{r}
describe(wine_data, na.rm = TRUE, IQR = TRUE) %>%
  dplyr::select(mean, sd, IQR, range, skew, kurtosis, se)
```

####Acid Index
The Acid Index variable is skewed to the right. The outliers are also high.
```{r}
par(mfrow=c(1,2))
acidIndex_density <- density(wine_data$AcidIndex)
plot(acidIndex_density, main="Acid Index")
polygon(acidIndex_density, col="red", border="red")

boxplot(wine_data$AcidIndex, main="Boxplot for Acid Index", col="blue")

x <- get_outliers(wine_data$AcidIndex)
x
```

####Alcohol
There does not appear to be any skew and seems to have a leptokurtic distribution.
```{r}
par(mfrow=c(1,2))
alcohol_density <- density(wine_data$Alcohol, na.rm=TRUE)
plot(alcohol_density, main="Alcohol")
polygon(alcohol_density, col="red", border="red")

boxplot(wine_data$Alcohol, main="Boxplot for Alcohol", col="blue")

x <- get_outliers(wine_data$Alcohol)
x
```

####Chlorides
COMMENTS
```{r}
par(mfrow=c(1,2))
chlorides_density <- density(wine_data$Chlorides, na.rm=TRUE)
plot(chlorides_density, main="Chlorides")
polygon(chlorides_density, col="red", border="red")

boxplot(wine_data$Chlorides, main="Boxplot of Chlorides", col="blue")

x <- get_outliers(wine_data$Chlorides)
x
```

####Citric Acid
COMMENTS
```{r}
par(mfrow=c(1,2))
citAcid_density <- density(wine_data$CitricAcid)
plot(citAcid_density, main="Citric Acid")
polygon(citAcid_density, col="red", border="red")

boxplot(wine_data$CitricAcid, main="Boxplot of Citric Acid", col="blue")

x <- get_outliers(wine_data$CitricAcid)
x
```

#### Density
COMMENTS
```{r}
par(mfrow=c(1,2))
dens_density <- density(wine_data$Density)
plot(dens_density, main="Density")
polygon(dens_density, col="red", border="red")

boxplot(wine_data$Density, main="Boxplot of Density", col="blue")

x <- get_outliers(wine_data$Density)
x
```

####Fixed Acidity
COMMENTS
```{r}
par(mfrow=c(1,2))
fixAcid_density <- density(wine_data$FixedAcidity)
plot(fixAcid_density, main="Fixed Acidity")
polygon(fixAcid_density, col="red", border="red")

boxplot(wine_data$FixedAcidity, main="Boxplot of Fixed Acidity", col="blue")

x <- get_outliers(wine_data$FixedAcidity)
x
```

####Free Sulfur Dioxide
COMMENTS
```{r}
par(mfrow=c(1,2))
freeSulf_density <- density(wine_data$FreeSulfurDioxide, na.rm=TRUE)
plot(freeSulf_density, main="Free Sulfur Dioxide")
polygon(freeSulf_density, col="red", border="red")

boxplot(wine_data$FreeSulfurDioxide, main="Boxplot of Free Sulfer Dioxide", col="blue")

x <- get_outliers(wine_data$FreeSulfurDioxide)
x
```

####Label Appeal
COMMENTS
```{r}
par(mfrow=c(1,2))
labelAppeal_density <- density(wine_data$LabelAppeal)
plot(labelAppeal_density, main="Label Appeal")
polygon(labelAppeal_density, col="red", border="red")

boxplot(wine_data$LabelAppeal, main="Boxplot for Label Appeal", col="blue")

x <- get_outliers(wine_data$LabelAppeal)
x
```

####pH
COMMENTS
```{r}
par(mfrow=c(1,2))
pH_density <- density(wine_data$pH, na.rm=TRUE)
plot(pH_density, main="pH")
polygon(pH_density, col="red", border="red")

boxplot(wine_data$pH, main="Boxplot for pH", col="blue")

x <- get_outliers(wine_data$pH)
x
```

####Residual Sugar
COMMENTS
```{r}
par(mfrow=c(1,2))
resSugar_density <- density(wine_data$ResidualSugar, na.rm=TRUE)
plot(resSugar_density, main="Residual Sugar")
polygon(resSugar_density, col="red", border="red")

boxplot(wine_data$ResidualSugar, main="Boxplot of Residual Sugar", col="blue")

x <- get_outliers(wine_data$ResidualSugar)
x
```

####Stars
COMMENTS
```{r}
par(mfrow=c(1,2))
stars_density <- density(wine_data$STARS, na.rm=TRUE)
plot(stars_density, main="Stars")
polygon(stars_density, col="red", border="red")

boxplot(wine_data$STARS, main="Boxplot for Stars", col="blue")

x <- get_outliers(wine_data$STARS)
x
```

####Sulphates
COMMENTS
```{r}
par(mfrow=c(1,2))
sulphate_density <- density(wine_data$Sulphates, na.rm=TRUE)
plot(dens_density, main="Sulphates")
polygon(dens_density, col="red", border="red")

boxplot(wine_data$Sulphates, main="Boxplot for Sulphates", col="blue")

x <- get_outliers(wine_data$Sulphates)
x
```

####TARGET
COMMENTS
```{r}
par(mfrow=c(1,2))
target_density <- density(wine_data$TARGET)
plot(target_density, main="Target Variable")
polygon(target_density, col="red", border="red")

boxplot(wine_data$FixedAcidity, main="Boxplot of the Target variable", col="blue")

x <- get_outliers(wine_data$TARGET)
x
```

####Total Sulfur Dioxide
COMMENTS
```{r}
par(mfrow=c(1,2))
totalSulf_density <- density(wine_data$TotalSulfurDioxide, na.rm=TRUE)
plot(totalSulf_density, main="Total Sulfur Dioxide")
polygon(totalSulf_density, col="red", border="red")

boxplot(wine_data$TotalSulfurDioxide, main="Boxplot of Total Sulfer Dioxide", col="blue")

x <- get_outliers(wine_data$TotalSulfurDioxide)
x
```

####Volatile Acidity
COMMENTS
```{r}
par(mfrow=c(1,2))
volAcid_density <- density(wine_data$VolatileAcidity)
plot(volAcid_density, main="Volatile Acidity")
polygon(volAcid_density, col="red", border="red")

boxplot(wine_data$VolatileAcidity, main="Boxplot of Volatile Acidity", col="blue")

x <- get_outliers(wine_data$VolatileAcidity)
x
```




##Correlation Analysis

The correlation matrix shown below highlights correlations among several predictor variables. Correlation between between `Target` and `Stars` is high at 0.6. The second highest correlation is between the `Target` variable and `Label Appeal` at 0.4.




```{r}
ggcorr(wine_data, method = "pairwise", label=TRUE, nbreaks=6)
```

Let us now look at the correlation between each response variable and the predictors.

```{r}
cor(wine_data$TARGET, wine_data$FixedAcidity, use="complete.obs")
cor(wine_data$TARGET, wine_data$VolatileAcidity, use="complete.obs")
cor(wine_data$TARGET, wine_data$CitricAcid, use="complete.obs")
cor(wine_data$TARGET, wine_data$ResidualSugar, use="complete.obs")
cor(wine_data$TARGET, wine_data$Chlorides, use="complete.obs")
cor(wine_data$TARGET, wine_data$FreeSulfurDioxide, use="complete.obs")
cor(wine_data$TARGET, wine_data$TotalSulfurDioxide, use="complete.obs")
cor(wine_data$TARGET, wine_data$Density, use="complete.obs")
cor(wine_data$TARGET, wine_data$pH, use="complete.obs")
cor(wine_data$TARGET, wine_data$Sulphates, use="complete.obs")
cor(wine_data$TARGET, wine_data$Alcohol, use="complete.obs")
cor(wine_data$TARGET, wine_data$LabelAppeal, use="complete.obs")
cor(wine_data$TARGET, wine_data$AcidIndex, use="complete.obs")
cor(wine_data$TARGET, wine_data$STARS, use="complete.obs")
```

###Multicollinearity

This section will test the predictor variables to determine if there is correlation among them. Variance inflaction factor (VIF) is used to detect multicollinearity, specifically among the entire set of predictors versus within pairs of variables.

Testing for collinearity among the predictor variables, we see that none of the numeric predictor variables appear to have a problem with collinearity based on their low VIF scores.
```{r}
numeric_fields <- dplyr::select_if(wine_data, is.numeric)[, 2:15]

usdm::vifcor(numeric_fields) 
```




#2. Data Preparation

##Missing Values

8 out of 15 variables contain missing values. The predictor `ResidualSugar`contains 616 (5%) missing values, `Chlorides` with 638 (5%), `FreeSulfurDioxide` with 647 (5%), `TotalSulfurDioxide` with 682 (5%), `pH` with 395 (3%), `Sulphates` with `1,210` (9%), `Alcohol` with 653 (5%), and `STARS` with 3,359 (26%) missing values .

The visualization of missing values below shows that missing values of `ResidualSugar`, `Chlorides` and YOJ are at 6 percent while INCOME is at 5 percent. The dataset was imputed to account for the missing values.

There are a total of 8,200 `NA` values within the 12,795 observations and 15 columns, a rate of **4.27%** of the entire data set.
```{r}
total_na <- sum(is.na(wine_data))
total_cells <- 12795 * 15
round(total_na *100 / total_cells,2)
```

```{r}
kable(sapply(wine_data, function(x) sum(is.na(x))))

plot_missing(wine_data, title="Wine Dataset - Missing Values (%)")
```

```{r warning=FALSE}
#Examine missing data pattern
#md.pattern(wine_data)

#Number of missing
aggr(wine_data, prop = T, numbers = T)
```


```{r}
df <- setNames(data.frame(colSums(wine_data==0, na.rm = T)), 'Count')
           
df$Variable <- rownames(df)

rownames(df) <- NULL

df %>% filter(!Variable %in% c("TARGET")) %>%  
ggplot(aes(x=reorder(Variable, Count), y=Count, fill=Count)) +
    geom_bar(stat="identity") + coord_flip() + guides(fill=FALSE) +
    xlab("Variable") + ylab("Number of 0 Values") + 
    ggtitle("Count of Zero Values by Variable") +
    geom_text(aes(label=Count), vjust=.5, hjust=-.1,position= position_dodge(width=0.5),size=3,  color="black")
```

##Assumptions of Missing Values

```{r}
#ggplot(wine_data,
#       aes(x = HOME_VAL,
#           y = JOB)) +
#geom_point() +
#    facet_wrap(~JOB) +
#    ggtitle("Home Value by Profession")

#Employment by income
#ggplot(wine_data,
#       aes(x = INCOME,
#           y = JOB)) +
#geom_point() +
#    facet_wrap(~JOB) +
#    ggtitle("Income Value by Profession")

#Replace student and missing HOME_VAL with 0
#insurance_train$INCOME <- as.numeric(str_replace_all(insurance_train$INCOME, "\\$|,", ""))
#insurance_train %>% mutate(HOME_VAL=replace(HOME_VAL, is.na(HOME_VAL) & JOB=="Student", 0)) -> insurance_train

#Replace Home Maker and missing Income with 0
#insurance_train$INCOME <- as.numeric(str_replace_all(insurance_train$INCOME, "\\$|,", ""))
#insurance_train %>% mutate(INCOME=replace(INCOME, is.na(INCOME) & JOB=="Home Maker", 0)) -> insurance_train

```


##Imput Recoded dataset

We will now impute the missing data on the recoded data set. None of the data to be imputed are categorical.

```{r}
#Process recoded data set
#insurance_train_impute <- mice(insurance_train,m=3,maxit=50,meth='cart',seed=500) 

#summary(insurance_train_impute)

#inspect imputed data
#densityplot(insurance_train_impute)

#Check imputed values
#insurance_train_impute$imp$AGE

#head(insurance_train_impute)

#get complete data ( 2nd out of 3)
#insurance_train_complete <- mice::complete(insurance_train_impute, 2)

#write.csv(insurance_train_complete, 'insurance_training_Impute.csv', row.names = F)

```

