---
title: "Assignment/HW5"
author: "Group 2"
date: "April 19, 2018"
output: html_document
---
Overview
In this homework assignment we explore, analyze, and model a data set containing information on approximately 12,000 commercially available wines. The variables are mostly related to the chemical properties of the wine being sold. The response variable is the number of sample cases of wine that were purchased by wine distribution companies after sampling a wine. These cases would be used to provide tasting samples to restaurants and wine stores around the United States. The more sample cases purchased, the more likely is a wine to be sold at a high end restaurant. 

A large wine manufacturer is studying the data in order to predict the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then it will be able to adjust their wine offering to maximize sales.

Our objective is to build a count regression model to predict the number of cases of wine that will be sold
given certain properties of the wine. 
```{r message=FALSE, warning=FALSE}
library(psych)
library(tidyverse)
library(knitr)
library(ggplot2)
library(reshape2)

```

```{r}
variable_names <- c("Index", "Target", "AcidIndex", "Alcohol", "Chlorides", 
                    "CitricAcid", "Density", "FixedAcidity", "FreeSulfurDioxide", "LabelAppeal", 
                    "ResidualSugar", "Stars", "Sulphates", "TotalSulfurDioxide", "VolatileAcidity", "pH")

description <- c("Identification Variable (do not use)", 
                   "Number of cases purchased", 
                   "Proprietary method of testing ottal acidity of wine by using a weighted average", 
                   "Alcohol content", 
                   "Chloride content of wine",
                   "Citric acid content",
                   "Density of wine",
                   "Fixed acidity of wine",
                   "Sulfur dioxide content of wine",
                   "Marketing score indidcating appeal of label design for consumers. Positive is like, negative is dislike",
                   "Resdiual sugar of wine",
                   "Wine rating by a team of experts. 4 stars = Excellent, 1 star = Poor",
                   "Sulfate content of wine",
                   "Total sulfur dioxide of wine",
                   "Volatile acid content of wine",
                   "pH of wine")

variable_type <-c("Predictor", "Response", "Predictor", "Predictor", "Predictor", 
                  "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", 
                  "Predictor", "Predictor", "Predictor",  "Predictor",  "Predictor",  "Predictor")

data_type <- c("quantitative", "quantitative", "quantitative", "quantitative", "quantitative", 
               "quantitative", "quantitative", "quantitative", "quantitative", "categorical", 
               "quantitative", "categorical", "quantitative", "quantitative", "quantitative", "quantitative")

wine_descriptions <- cbind.data.frame(variable_names, description, variable_type, data_type)

colnames(wine_descriptions) <- c("Variable Name", "Description", "Variable Type", "Data Type") 

kable(wine_descriptions)
```



The data were loaded into a Github repository for better reproducability. 
```{r}
wine_data <- read.csv("https://raw.githubusercontent.com/621-Group2/HW5/master/wine-training-data.csv")
class(wine_data)
```

##1. DATA EXPLORATION

The training set has 12,795 observations on 16 variables, including the Target (response) variable.
```{r}
glimpse(wine_data)
```
After taking a quick look at the head and tail of the data, the first thing noticed was the `INDEX` column is not in perfect sequence. There are also a number of NA's we will have to deal with. 
```{r}
head(wine_data)

tail(wine_data)
```
A simple summary gives us a good overview. As noticed in the head and tail, there are plenty of NA's. As much as 26% of the data is missing from the `STARS` column. Seems like the "experts" are afraid to weigh in!
```{r}
summary(wine_data)
```

Looking at the skewness coefficient and kurtosis, `AcidIndex` appears to be skewed.
```{r}
describe(wine_data, na.rm = TRUE, IQR = TRUE) %>% 
  select(mean, sd, IQR, range, skew, kurtosis, se)
```

We need a visual representation of these numbers in order to glean more insight. Density and box plots will be the first method.
```{r message=FALSE, warning=FALSE}

#Density
par(mfrow = c(3, 3))
wine_data %>%  
  melt %>% 
  ggplot(aes(x= value)) + 
    geom_density(fill='maroon') + facet_wrap(~variable, scales = 'free')
  

#Boxplot
wine_data %>%
  melt %>%
  ggplot(mapping = aes(x= "", y = value)) + 
    geom_boxplot(fill="maroon") + facet_wrap(~variable, scales = 'free')
```
Our suspicions are confirmed. `AcidIndex` appears to be highly skewed left, with a range of 4 to 17. 

```{r warning=FALSE}
library(PerformanceAnalytics)
par(mfrow = c(2,1))
chart.Correlation(wine_data[1:5])
chart.Correlation(wine_data[6:10])
chart.Correlation(wine_data[11:16])
#library(psych)
#pairs.panels(wine_data[2:16])

#library(corrplot)
#x <- cor(wine_data[1:16])
#corrplot(x)
```








